Summer 2025 ToDo

### 2025-05-30 Relations/Paper/Formalization.tex
Formatting updates throughout. Updated references to agda code.

\section{Formalization of \terese}
\label{sec:Formalization}
This sections is a guide to our formalization of \terese and explains the deviations we take from the text.

\subsection{Formalization of Definitions}\label{subsec:def}
The closure operators are formalized in \texttt{ARS-Closure.agda}. 
The definitions grouped as confluent in \terese are formalized in \texttt{ARS-Base.agda}. 
The definitions grouped as normalizing in \terese are formalized in \texttt{ARS-Properties.agda}.

@@

\subsection{Propositions}
There are three sets of propositions given in \terese: \texttt{1.1.9}, \texttt{1.1.10}, \texttt{1.1.11}. 
The formalization of the proofs are in \texttt{ARS-Propositions.agda}. These formalizations are straightforward 
and as such are not the focus of this paper.

@@

($i$) is formalized in:
\begin{itemize}
    \item \verb|CR→NP : R isCR → R isNP₌|
    \item \verb|NP→UN : R isNP₌ → R isUN|
\end{itemize}

($ii$) is formalized also from the weaker hypothesis of $\gUNto$:
\begin{itemize}
    \item \verb|ii : R isWN × R isUN → R isCR|
    \item \verb|UN→∧WN→CR : R isUN→ → R isWN → R isCR|
\end{itemize}

($iii$) is formalized in:
\begin{itemize}
    \item \verb|iii : subcommutative R → R isCR|
\end{itemize}

\subsubsection{Theorem 1.2.3}
The following definition is required for this theorem.
\begin{definition}\hfill
    \begin{description}
        \item[\ule{$a \in CP_R$}] \emph{$a$ has the cofinality property} if there exists a 
        (not necessarily strictly) increasing sequence starting from $a$ such that every
        reduct of $a$ reduces to some element of the sequence.
    \end{description}
\end{definition}

There are four parts to this theorem. 
\begin{description}
    \item[($i$)] $\gWN \land \gUN \implies \gBP$
    \item[($ii$)] $\gBP \land \gInc \implies \gSN$  
    \item[($iii$)] $\gWCR \land \gWN \land \gInc \implies \gSN$
    \item[($iv$)] $\gCR \iff \gCP$ ($\gCR \implies \gCP$ only holds if the ARS is countable.)
\end{description}

($i$) is formalized in: 
\begin{itemize}
    \item \verb|i : R isWN → R isUN → R isBP|
    \item \verb|i→ : R isWN → R isUN→ → R isBP|
\end{itemize}

In $\mathtt{i\to}$ we again provide a slight improvement by using $\gUNto$ rather than $\gUN$.

A formalization of ($ii$) is provided in: 
\begin{itemize}
    \item \verb|iiSeq : R isWN → R isUN → R isRP → isWFseq- (~R R)|
\end{itemize}
This is an inherently classical result, and so the result is only classically equivalent to $\gSN$.

($iii$) is formalized in: 
\begin{itemize}
    \item \verb|iii : R isWN → R isWCR → R isRP- → dec (SN) → R isSN|
\end{itemize}
As can be seen we have substituted $\gInc$ with $\gRPm$ which generalizes this theorem. We have also included the assumption that $\gSN$ 
is decidable as this property enables us 
to construct a witness for an element that either does or does not have the property $\SN$. Such a witness was necessary for our proof. 
% \sacomment{Link to decidable in later section?}

The reverse implication of ($iv$) is formalized in: 
\begin{itemize}
    \item \verb|iv : R isCP → R isCR|
\end{itemize}
We did not formalize $\gCR \to \gCP$ as it requires cardinality assumptions.

% the ARS to be countable.

>>>>

\section{Formalization of \terese}
\label{sec:Formalization}
This sections is a guide to our formalization of \terese $\,$ and explains the deviations we take from the text.

\subsection{Formalization of Definitions}\label{subsec:def}
The closure operators are formalized in \texttt{Closure.agda}. 
The definitions grouped as confluent in \terese $\,$ are formalized in \texttt{Base.agda}. 
The definitions grouped as normalizing in \terese $\,$ are formalized in \texttt{Properties.agda}.

@@

\subsection{Propositions}
There are three sets of propositions given in \terese: \texttt{1.1.9}, \texttt{1.1.10}, \texttt{1.1.11}. 
The formalization of the proofs are in \texttt{Propositions.agda}. These formalizations are straightforward 
and as such are not the focus of this paper.

@@

\begin{proof}
    See, \\
    \verb|CR→NP : R isCR → R isNP₌|\\
    \verb|NP→UN : R isNP₌ → R isUN|\\
    \verb|ii : R isWN × R isUN → R isCR|\\
    \verb|UN→∧WN→CR : R isUN→ → R isWN → R isCR|\\
    \verb|iii : subcommutative R → R isCR|\\
    in \texttt{Theorems.agda}.
\end{proof}
\begin{remark}
    In $\mathtt{UN→∧WN→CR}$ a more general proof of ($ii$) is given using $\UNto$.
\end{remark}

\subsubsection{Theorem 1.2.3}
The following definition is required for this theorem.
\begin{definition}\hfill
    \begin{description}
        \item[\ule{$a \in CP_R$}] $a$ has the \emph{cofinality property} if there exists an 
        (not necessarily strictly) increasing sequence starting from $a$ such that every
        reduct of $a$ reduces to some element of the sequence.
    \end{description}
\end{definition}

There are four parts to this theorem. 
\begin{description}
    \item[($i$)] $\gWN \land \gUN \implies \gBP$
    \item[($ii$)] $\gBP \land \gInc \implies \gSN$  
    \item[($iii$)] $\gWCR \land \gWN \land \gInc \implies \gSN$
    \item[($iv$)] $\gCR \iff \gCP$ ($\gCR \implies \gCP$ only holds if the ARS is countable.)
\end{description}
\begin{proof}
    See, \\
    \verb|i : R isWN → R isUN → R isBP|\\
    \verb|i→ : R isWN → R isUN→ → R isBP|\\
    \verb|iiSeq : R isWN → R isUN → R isRP → isWFseq- (~R R)|\\
    \verb|iii : R isWN → R isWCR → R isRP- → dec (SN) → R isSN|\\
    \verb|iv : R isCP → R isCR| \\
    in \texttt{Theroems.agda}.
\end{proof}

\begin{remark} \hfill
    \begin{itemize}        
        \item In $\mathtt{i\to}$ we again provide a slight improvement by using $\gUNto$ rather than $\gUN$.
        \item ($ii$) is inherently classical. In $\mathtt{iiSeq}$, we show that these properties imply $\isWFseqm$; establishing the implication to $\gSN$ would require additional classical assumptions. The definition and further discussion of $\isWFseqm$ are provided in Section~\ref{sec:Well-foundedness}. 
        \item In $\mathtt{iii}$, we replace $\gInc$ with the more general property $\gRPm$. We also assume that $\gSN$ is decidable, as this allows the construction of a witness for whether an element possesses the property $\SN$, which is essential for our proof.
        \item We did not formalize $\gCR \to \gCP$ as it requires cardinality assumptions.
    \end{itemize}
\end{remark}

### 2025-05-30 Relations/Paper/Implications.tex
Formatting updates throughout. Updated references to agda code.

As discussed in Subsection \ref{subsec:def} we put forward $\RP$ as an alternative for $\Inc$. It is through examining $\RP$ that 
we came to the 'minimal` family of properties below. These properties help to augment our ARS hierarchy.

@@

Trivially, the following holds. 
\begin{proposition}\label{prop:nftomf}
    $\NF \implies \MF$ 
\end{proposition}    
\begin{proof}
    See, \verb|NF ⊆ MF : ∀x → NFx → MFx| in \texttt{ARS-Implications.agda}.
\end{proof}

$\SMseq$ is only classically equivalent to $\SM$. The 
following proposition relates these new definitions with our previous $\RP$ definitions.
\begin{proposition}\label{prop:SMRP}
    $SMseq \leftrightarrow RP \land BP$
\end{proposition}
\begin{proof}
    See,\\ \verb|RP∧BP→SMseq : R isRP → R isBP → ∀ {x : A} → SMseq R x| \\ \verb|RisSMseq→RisRP : (∀ {x : A} → SMseq R x) → R isRP|
    \\ \verb|RisSMseq→RisBP : (∀ {x : A} → SMseq R x) → R isBP| in \texttt{ARS-Implications.agda}.
\end{proof}

@@

The implications in this diagram are formalized in \texttt{ARS-Implications.agda}. 
The hierarchy for confluencce requires no classical assumptions, however this is not the case for the hierarchy of terminating properties.
The implication from $\SN$ to $\WN$ relies on the classical decidability of the property 
of being $R$-minimal, as demonstrated in our formalization:

>>>>

As discussed in Subsection \ref{subsec:def} we put forward $\RP$ as an alternative for $\Inc$. It is through examining $\RP$ that 
we came to the `minimal' family of properties below. These properties help to augment our ARS hierarchy.

@@

Trivially, the following holds. 
\begin{proposition}\label{prop:nftomf}
    $\NF \implies \MF$ 
\end{proposition}    
\begin{proof}
    See, \verb|NF ⊆ MF : ∀x → NFx → MFx| in \texttt{Implications.agda}.
\end{proof}

$\SMseq$ is only classically equivalent to $\SM$. The 
following proposition relates these new definitions with our previous $\RP$ definitions.
\begin{proposition}\label{prop:SMRP}
    $SMseq \leftrightarrow RP \land BP$
\end{proposition}
\begin{proof}
    See,\\ 
    \verb|RP∧BP→SMseq : R isRP → R isBP → ∀ {x : A} → SMseq R x| \\ 
    \verb|RisSMseq→RisRP : (∀ {x : A} → SMseq R x) → R isRP|\\ 
    \verb|RisSMseq→RisBP : (∀ {x : A} → SMseq R x) → R isBP| \\
    in \texttt{Implications.agda}.
\end{proof}

@@

The implications in this diagram are formalized in \texttt{Implications.agda}. 
The confluence hierarchy requires no classical assumptions, however this is not the case for the hierarchy of terminating properties.
The implication from $\SN$ to $\WN$ relies on the classical decidability of the property 
of being $R$-minimal, as demonstrated in our formalization:

### 2025-05-30 Relations/Paper/Summer_2025_TODO.tex
Updated the todos
Formatting updates throughout. Updated references to agda code.

    \item (Review 1) Improve the layout on line 270 where we set out iiSeq. This should get fixed with an overall improvement in how we set out definitions. We should stop using itemize and verb combination.

@@

    \item (Review 2) Section 3 is "full of repeitions and paraphrases between the statements in natural language and corresponding formalization". We can try and think of a way to get it to flow bettwe, we do want to provide natural language summaries of what is in the code!
    \item (Review 2) Struggled to understand the implications tables. We should add a reminder that bold font is global. We should rethink the tables generally to make them clearer. Don't comma seperate and don't denote counterexamples as CE-XX.

@@

\section*{Week starting April 28th}
\begin{itemize}
  \item Below notes from meeting on April 25th.
  \begin{todolist}
    \item Fix Wellfounded.agda after changes made during meeting.
    \item Carry on refactoring Wellfounded.agda once fixed, breaking it up if necessary.
    \item Think about how to organize the Git repo going forward. For example, could be good to have a parent folder with ARS and WF, with a private dir for paper 
    \item (Research item) How to go from WFseq to WFacc - possible that the Kleene Tree is a solution.
    \item (Research item) Go deeper in the related works and see which are most relevant. Make a list of works and give summaries to be used in future papers. 
    \item (Research item) Find a concrete application of generalized Newman's Lemma. 
    \item (Research item) Use TeReSe to create a list of rewrite systems.
    \item (Research item) Work on the missing implicaitons from the paper. So far we believe this is just CR $\to$ CP in the ARS section. 
    \item (Research item) Formalize a first order term rewriting system (I can't remember what the motivation for this was).
    \item (Research item) Carry out research on rewriting counterexamples like our own so we can try and put them into some sort of context. Have our counerexamples been shown before?
    \item (Research item) Do we want to reincorporate Knaster Tarski?  
  \end{todolist}
\end{itemize}
\end{document}

>>>>

    \item [\done] (Review 1) Improve the layout on line 270 where we set out iiSeq. This should get fixed with an overall improvement in how we set out definitions. We should stop using itemize and verb combination.

@@

    \item (Review 2) Section 3 is "full of repetitions and paraphrases between the statements in natural language and corresponding formalization". We can try and think of a way to get it to flow better, we do want to provide natural language summaries of what is in the code!
    \item [\done] (Review 2) Struggled to understand the implications tables. We should add a reminder that bold font is global. We should rethink the tables generally to make them clearer. Don't comma seperate and don't denote counterexamples as CE-XX.

@@

\section*{Week starting April 28th}
\begin{itemize}
  \item Below notes from meeting on April 25th.
  \begin{todolist}
    \item Fix Wellfounded.agda after changes made during meeting.
    \item Carry on refactoring Wellfounded.agda once fixed, breaking it up if necessary.
    \item Think about how to organize the Git repo going forward. For example, could be good to have a parent folder with ARS and WF, with a private dir for paper 
    \item (Research item) How to go from WFseq to WFacc - possible that the Kleene Tree is a solution.
    \item (Research item) Go deeper in the related works and see which are most relevant. Make a list of works and give summaries to be used in future papers. 
    \item (Research item) Find a concrete application of generalized Newman's Lemma. 
    \item (Research item) Use TeReSe to create a list of rewrite systems.
    \item (Research item) Work on the missing implicaitons from the paper. So far we believe this is just CR $\to$ CP in the ARS section. 
    \item (Research item) Formalize a first order term rewriting system (I can't remember what the motivation for this was).
    \item (Research item) Carry out research on rewriting counterexamples like our own so we can try and put them into some sort of context. Have our counerexamples been shown before?
    \item (Research item) Do we want to reincorporate Knaster Tarski?  
  \end{todolist}

\end{itemize}
\section*{Week starting 26th May}
\begin{todolist}
  \item Double check the counterexamples match in the table to the diagrams.
\end{todolist}
\end{document}
